---
title: "Landfire Apoidea Masking"
author: "Nicolas J. Dowdy"
date: "2021-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(12345) # set seed for code reproducibility
# set resource locations
setwd("scripts/")
landfire_input = "../input/LF_140EVT.csv"
# dblbnd.adf (aka bnd.adf) -- The boundary file contains the minimum and maximum x,y coordinates for an ESRI_grid. The boundary is a rectangle that encompasses the cells of a grid; it is stored in map coordinates. All grid BNDs are stored in double precision. This file will have a corresponding .nit and .data files in the INFO directory of an ESRI_grid. This information is stored in INFO format. https://www.loc.gov/preservation/digital/formats/fdd/fdd000281.shtml
boundary_grid = "../input/dblbnd.adf"
landfire_shapes = "../input/extracted/landfire-files"
raster_location = "../input/extracted/range-data-rasters"
shape_files_location = "../input/extracted/shape-files"
# set variables
abundant_cutoff = 7500
```

## Procedure

1. Get landfire classification for each observation
2. Use dplyr to count the landfire classes aggregated by species
3. For each species, limit the landfire classes to those above 10% and sum the rasters together
4. Mask the range raster for each species by the landfire raster from step #3
5. Save the resulting range raster for each species
6. Repeat for different resolutions
7. Profit.

```{r}
# load libraries for this chunk
library(tidyr)
library(dplyr)
library(raster) # install libxt-dev
library(rgdal) # install libgdal-dev
library(rgeos) # install libgeos-dev
library(stringr)

# function for generating paths to raster files
path2file <- function(dataset=NULL, parent=NULL, type=NULL) {
  if(is.null(parent)){
    print('Provide valid parent.')
  } else if(is.null(dataset)){
    print('Provide valid dataset.')
  } else if(!(parent %in% c('usa', 'landfire'))) {
    print('Provide valid parent.')
  } else {
    if(parent == 'landfire') {
        parent = landfire_shapes
        if(is.null(type)){
          dataset = dataset
        } else if(type == 'grid') {
          dataset = paste(dataset,'_raster.grd',sep='')
        }
    } else if(parent == 'usa') {
        parent = shape_files_location
    }
    file.path(parent, dataset)
  }
}

# load landfire metadata
landfire = read.csv(landfire_input)

# define and load shape files for each landfire type being considered
barren_shape <- readOGR(path2file('barren', 'landfire'))
sparse_shape <- readOGR(path2file('sparse', 'landfire'))
tree_shape <- readOGR(path2file('tree', 'landfire'))
herb_shape <- readOGR(path2file('herb', 'landfire'))
shrub_shape <- readOGR(path2file('shrub', 'landfire'))
agriculture_shape <- readOGR(path2file('agriculture', 'landfire'))
developed_shape <- readOGR(path2file('developed', 'landfire'))

# define and load raster grids
barren_grid <- raster(path2file('barren', 'landfire', 'grid'))
sparse_grid <- raster(path2file('sparse', 'landfire', 'grid'))
tree_grid <- raster(path2file('tree', 'landfire', 'grid'))
herb_grid <- raster(path2file('herb', 'landfire', 'grid'))
shrub_grid <- raster(path2file('shrub', 'landfire', 'grid'))
agriculture_grid <- raster(path2file('agriculture', 'landfire', 'grid'))
developed_grid <- raster(path2file('developed', 'landfire', 'grid'))

# TODO: ensure raster grids are compatible with north american / range grids


```


<!-- r=raster("dblbnd.adf") -->
<!-- Richness=spTransform(richness,CRS(proj4string(r))) -->
<!-- values <- extract(r,richness) -->
<!-- evt <- cbind.data.frame(coordinates(richness),values) -->

<!-- you will then want to match with the evt key (I'll also put that in dropbox) and then you'll want to use the "EVT_LF" column to filter and figure out what land types to use for that species based on how many times the land type is present. -->

```{r}
## read in the shape files and crop
usa=readOGR(path2file(parent='usa', dataset='usa'))
globe=readOGR(path2file(parent='usa', dataset='continents'))
NAm=globe[globe$CONTINENT=='North America',]
NAm=crop(NAm,extent(-165,-60,8,85))
usaWGS=spTransform(usa,CRS(proj4string(NAm)))

# read in occurrence data
test <- read.csv('../input/test_df_small.csv')

# fix scientificName formatting
test$scientificName <- gsub(' ', '_', test$scientificName)

# make a lat/long combo for grabbing unique values
test$longlat <- paste(test$finalLongitude, test$finalLatitude, sep='_')
  
# define taxa
taxa <- unique(test$scientificName) # get a list of unique taxa
grd_files <- list.files(raster_location, pattern = "\\.grd$") # generate a list of raster locations
name <- c() # allocate as a temporary list

taxon = taxa[1] # testing

for(taxon in taxa) {
  raster_file = grd_files[grep(taxon, grd_files)] # find the taxon raster file

  # load range map
  taxon_raster = raster(paste(raster_location, raster_file, sep="/")) # load the raster
  raster_usawgs = mask(taxon_raster, usaWGS) # mask the raster to the USA shapefile
  raster_nam = mask(taxon_raster, NAm) # mask the raster to the North America shapefile
  plot(raster_nam) # plot the raster as sanity check
  points(x=points$long, y=points$lat, pch=16, size=3)
  
  # get unique lat/long values for taxon
  taxon_test <- test[which(test$scientificName == taxon),]
  points <- as.data.frame(unique(taxon_test$longlat))
  colnames(points) <- 'longlat'
  points <- as.data.frame(str_split_fixed(points$longlat, "_", 2))
  colnames(points) <- c('long', 'lat')
  points$lat <- as.numeric(points$lat)
  points$long <- as.numeric(points$long)
  
  # assign landfire types to occurrences
  occurrences <- points
  barren<-cbind(raster::extract(barren_grid, occurrences, df = T), occurrences)$layer
  sparse<-cbind(raster::extract(sparse_grid, occurrences, df = T), occurrences)$layer
  tree<-cbind(raster::extract(tree_grid, occurrences, df = T), occurrences)$layer
  herb<-cbind(raster::extract(herb_grid, occurrences, df = T), occurrences)$layer
  shrub<-cbind(raster::extract(shrub_grid, occurrences, df = T), occurrences)$layer
  agriculture<-cbind(raster::extract(agriculture_grid, occurrences, df = T), occurrences)$layer
  developed<-cbind(raster::extract(developed_grid, occurrences, df = T), occurrences)$layer
  occurrences$id <- as.numeric(rownames(occurrences))
  occurrence_types <- cbind(occurrences, barren, sparse, tree, herb, shrub, agriculture, developed)
  occurrence_types$none <- abs(rowSums(occurrence_types[,4:10], na.rm=TRUE)-1)
  
  occurrences_fixed <- occurrence_types %>%
                       group_by(id, lat, long) %>%
                       gather(landfire_type, val, -id, -lat, -long) %>%
                       filter(val == 1) %>%
                       dplyr::select(-val) %>%
                       arrange(id)
  
  occurrences_fixed$longlat <- paste(occurrences_fixed$long, occurrences_fixed$lat, sep='_')
  taxon_test_merged <- merge(occurrences_fixed, taxon_test, by = 'longlat')
  
  # count up occurrences within each landfire type
  landfire_count <- taxon_test_merged %>%
  group_by(landfire_type) %>%
  summarise(count = n() ) %>%
  mutate( prop = 100*(count / sum(count)) )
  
  # determine cutoff for landfire inclusion
  if(sum(landfire_count$count)>abundant_cutoff){
    landfire_cutoff = 5
  } else {
    landfire_cutoff = 10
  }
  
  # determine which landfire types meet cutoff
  landfire_keepers <- landfire_count[which(landfire_count$prop>5),]$landfire_type
  
  # TODO: stack() grids of landfire_keepers
  
  # TODO: mask() taxon range with the landfire_keepers
  
}

```

